A Study of the Effectiveness of Usage Examples in REST API Documentation

### 2017-03-20 controlled_study.tex
better wording
caption
wording
low fat diet for related work section

We selected the WordPress REST API V2 for this study. WordPress is a blog-like open-source (R2) framework used by over 409 million people to visit 23.6 billion pages each month. The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, tags, etc.

The WordPress REST API V2 has been published and maintained since May 2015. Before January 2017, the WordPress REST API V2 was distributed as a plug-in where WordPress users could optionally install the API component. The following statistics are for the plug-in installation numbers between May 2015 and October 2016:

@@

\begin{subfigure}[t]{4in}
\centering
\includegraphics[width=4in]{original.png}
\caption{Original API Documentation}
\label{fig:original}
\end{subfigure}
~
\begin{subfigure}[t]{3in}
\centering
\includegraphics[width=2.7in]{forked.png}
\caption{Usage Example from the Enhanced API Documentation}
\label{fig:enhanced}
\end{subfigure}
\caption{Screenshots of Original vs. Enhanced API Documentation}
\end{figure*}

@@

Based on the lessons learned from the pilot studies, we designed the actual study process as follows: the original WordPress REST API documentation was enhanced and a total of 3 API usage examples were added to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post. Figure \ref{fig:original} shows a screenshot of the original API documentation related to T4 where the API client developers are provided with a reference table describing the different properties that can be used to create a $Post$ object\footnote{\url{http://v2.wp-api.org/reference/posts/}}. Figure \ref{fig:enhanced} shows a screenshot of the enhanced API documentation with a cURL based usage example. cURL is used because it's used by elsewhere in the original API documentation. In the enhanced API documentation, the example shows one possible API call with realistic values for the data that is described in the reference table and associated API response headers and body.

@@

\subsubsection{API Learning Obstacles}
  Researchers have published several papers on the topic of API learning obstacles. Robillard et al. used a mixed approach of surveys and in-person interviews with professional software engineers to understand the API learning obstacles \cite{Robillard_what_makes} \cite{Robillard_a_field_study}. They found that many obstacles for API learnabilty were related to API documentation and five impactful factors for API documentation as follows: documentation of intent, code examples, matching APIs with scenarios, penetrability of the API, and format and presentation. Uddin and Robillard surveyed 323 professional software engineers to understand the different ways in which a piece of documentation can be unfit for purpose \cite{g_uddin}. They identified ambiguity, incompleteness, and incorrectness as the three severest problems that lead API documentations to fail to answer the information needs of developers. Duala-Ekoko and Robillard performed a study with twenty programmers to understand the types of questions that developers have while using an unfamiliar API \cite{Duala-Ekoko:2012:AAQ:2337223.2337255}. They categorized the API related information needs into twenty generic questions that can be used to analyze the quality of API documentation. Myers et al. performed a study of enterprise SOA API usability and found that the study participants had limited success finding the relevant API elements to perform an API task \cite{Myers_study}. Among other factors, they recommended providing code examples in the API documentation to help API client developers with sample solutions to different patterns of API tasks.

Our work fits in the general space of studies of API documentation obstacles by focusing specifically on problems related to the absence of examples, and by expanding into the study of REST APIs.

\subsubsection{Crowd-Sourced API Usage Examples}
Prior research investigated the sourcing API usage examples from the Internet. Wang et al. performed an exploratory study to understand the current state of API related knowledge available on the Internet that can be leveraged by API client developers \cite{6462686}. They searched the web for usage examples related to the API of five popular Java libraries and found that on an average API examples could be found for 77\% of the 4,637 APIs included in the selected libraries. Moreover, they found that the crowd-sourcing sites accounted for 93.7\% of the usage examples compared to 6.13\% that were published on the official documentation. An evaluation of Jadite showed that developers were three times faster to complete API tasks when given placeholder API elements in combination with usage examples sourced from the web \cite{5295283}. Nasehi et al. analyzed StackOverflow threads to understand the characteristics of good and bad code examples based on user provided votes \cite{Nasehi_what_makes}. They found that commonly down-voted API related answers lacked code examples, explanations, and shortcoming of solutions. They recommended API developers to evolve the documentation with usage examples to answer API client developer questions that were not anticipated in the existing documentation. Treude et al. proposed a machine learning based approach to find relevant API documentation from StackOverflow using both the text and metadata found from the questions and answers~\cite{Treude:2016:AAD:2884781.2884800}. Kim et al. presented a technique to automatically augment code examples from the web using code search tools to Java APIs \cite{Kim:2009:AEJ:1747491.1747552}. Jiau et al. observed a severe inequality within the context of crowd-sourced API usage examples where most of the content were related to the popular APIs \cite{jiau}.

Existing research on crowd-sourced API documentation has mostly focused finding crowd-sourced usage examples of local APIs. In this paper, we focused on the impact of usage examples for REST APIs, where the API client development programming language is agnostic to the language used to implement the API. This important distinction with local APIs makes it hard to leverage the techniques proposed in the aforementioned papers to find relevant REST API usage examples.

\subsubsection{Measuring API Usability}
Several papers have been published on the topic of measuring API usability. Rama et al. presented a set of formulas for computing a measure of API usability based on the API's structural components such as classes, methods, parameters, return values, thread-safety, etc~\cite{rama}. Scheller et al. presented a framework for automatically measuring the complexity of an API~\cite{scheller2015automated}. They identified a list of measurable API properties and provided formulas to compute complexities related to the interfaces, implementation and setup of an API. Grill et al. proposed an HCI based approach that can be used by API developers to understand and improve on the problem areas related to an API \cite{Grill2012}. They suggested using a combination of expert opinion and developer workshops on APIs to identify and collect feedback about API and its documentation related problems. To improve API usability, Farooq et al. proposed using peer reviews of API code in addition to API usability studies to uncover API related bugs and incorporate feedback regarding the API elements~\cite{Farooq:2010:AUP:1753326.1753677}. In our work, instead of measuring overall API usability, we have focused specifically on the relationship of REST API usability with respect to usage examples.

\subsubsection{Controlled Studies}
Several researchers published the results of controlled studies on the topic of API usability. Nasehi et al. performed a controlled study to understand if API unit tests can be used to provide as usage examples to facilitate API client developers \cite{5609553}. They grouped participants into two groups, and one of the groups was provided with API unit tests in addition to the documentation. The two groups performed the same set of API tasks. The researchers found that the examples from the unit tests helped understanding the API concepts better but it was challenging for the participants to locate relevant examples from the corpus of unit tests. They recommended automated extraction of relevant high-level API usage examples from API unit tests. Endrikat performed a controlled study with four groups of participants to understand the impact of API documentation on APIs that are implemented using programming languages with static and dynamic type systems \cite{Endrikat:2014:ADS:2568225.2568299}. Participants were given a set of failing unit tests and were asked to make them pass by writing code using the studied API. They defined the API client developer productivity in terms of the time to get the tests passing. They found that the participant group using a statically typed API with explicit documentation were more productive than the other groups. Ellis et al. performed a controlled study with two groups of Java developers to understand the impact of using a constructor vs. a factory method design pattern on API client development time \cite{Ellis:2007:FPA:1248820.1248863}. They found that API client developers needed more time to complete API tasks when the API requires the use a factory method to instantiate objects compared to using a constructor.

Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned API related controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of a local API; and, the participants are professional software engineers.

>>>>

We selected the WordPress REST API V2 for this study. WordPress is a blog-like open-source (R2) framework used by over 409 million people to visit 23.6 billion pages each month \footnote{\url{https://wordpress.com/activity/}}. The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, tags, etc.

The WordPress REST API V2 has been published and maintained since May 2015. Before January 2017, the WordPress REST API V2 was distributed as a plug-in where WordPress users could optionally install the API component. The following statistics are for the plug-in installation numbers between May 2015 and October 2016:\footnote{\url{https://wordpress.org/plugins/rest-api/stats/}}

@@

\begin{subfigure}[t]{4in}
\centering
\includegraphics[width=4in]{original.png}
\caption{Original API Documentation}
\label{fig:original}
\end{subfigure}
~
\begin{subfigure}[t]{3in}
\centering
\includegraphics[width=2.7in]{forked.png}
\caption{API Documentation Enhanced with an Example}
\label{fig:enhanced}
\end{subfigure}
\caption{Screenshots of Original vs. Enhanced API Documentation}
\end{figure*}

@@

We designed the actual study protocol as follows: we enhanced the original WordPress REST API documentation and added a total of 3 API usage examples to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post. We used the WordPress API unit tests to find relevant data for these examples. Figure \ref{fig:original} shows a screenshot of the original API documentation related to T4 where the API client developers are provided with a reference table describing the different properties that can be used to create a $Post$ object\footnote{\url{http://v2.wp-api.org/reference/posts/}}. Figure \ref{fig:enhanced} shows a screenshot of the enhanced API documentation with a cURL based usage example. cURL is used because it's used by elsewhere in the original API documentation. In the enhanced API documentation, the example shows one possible API call with realistic values for the data that is described in the reference table and associated API response headers and body.

@@

\subsubsection{API Learning Obstacles}
Robillard et al. used a mixed approach of surveys and in-person interviews with professional software engineers to understand the API learning obstacles \cite{Robillard_what_makes} \cite{Robillard_a_field_study}. They found that many obstacles for API learnabilty were related to API documentation. They identified five impactful factors for API documentation as follows: documentation of intent, code examples, matching APIs with scenarios, penetrability of the API, and format and presentation. Uddin and Robillard surveyed 323 professional software engineers to understand the different ways in which a piece of documentation can be unfit for purpose \cite{g_uddin}. They identified ambiguity, incompleteness, and incorrectness as the three severest problems that lead API documentations to fail to answer the information needs of developers. Duala-Ekoko and Robillard performed a study with 20 programmers to understand the types of questions that developers have while using an unfamiliar API \cite{Duala-Ekoko:2012:AAQ:2337223.2337255}. They categorized the API related information needs into twenty generic questions that can be used to analyze the quality of API documentation. Myers et al. found that participants had limited success finding the relevant API elements to perform an API task using an enterprise API \cite{Myers_study}. Among other factors, they recommended providing code examples in the API documentation to help API client developers. Our work fits in the general space of studies of API documentation obstacles by focusing specifically on problems related to the absence of examples, and by expanding into the study of REST APIs.

\subsubsection{Crowd-Sourced API Usage Examples}
Wang et al. performed an exploratory study to understand the current state of API related knowledge available on the Internet \cite{6462686}. They searched the web for usage examples related to the API of five popular Java libraries and found that on an average API examples could be found for 77\% of the 4,637 APIs included in the selected libraries. Moreover, they found that the crowd-sourcing sites accounted for 93.7\% of the usage examples compared to 6.13\% that were published on the official documentation. An evaluation of Jadite showed that developers were three times faster to complete API tasks with access to auto-referred crowd-sourced usage examples~\cite{5295283}. Nasehi et al. analyzed StackOverflow threads to understand the characteristics of good and bad code examples based on user provided votes \cite{Nasehi_what_makes}. They found that commonly down-voted API related answers lacked code examples, explanations, and shortcoming of solutions. Treude et al. proposed a machine learning based approach to find relevant API documentation from StackOverflow using both the text and metadata found from the questions and answers~\cite{Treude:2016:AAD:2884781.2884800}. Kim et al. presented a technique to automatically augment code examples from the web using code search tools to Java APIs \cite{Kim:2009:AEJ:1747491.1747552}. Jiau et al. observed a severe inequality within the context of crowd-sourced API usage examples where most of the content were related to the popular APIs \cite{jiau}. Existing research on crowd-sourced API documentation has mostly focused finding usage examples of local APIs. In this paper, we focused on REST APIs, where the API client development programming language is agnostic to API implementation language. This distinction with local APIs makes it hard to leverage the techniques proposed in the aforementioned papers for finding crowd-sourced REST API usage examples.

\subsubsection{Measuring API Usability}
Rama et al. presented a set of formulas for computing a measure of API usability based on the API's structural components such as classes, methods, parameters, return values, thread-safety, etc~\cite{rama}. Scheller et al. presented a framework for automatically measuring the complexity of an API~\cite{scheller2015automated}. They identified a list of measurable API properties and provided formulas to compute complexities related to the interfaces, implementation and setup of an API. Grill et al. suggested using a combination of expert opinion and developer workshops to identify and collect feedback about API and its documentation related problems \cite{Grill2012}. Farooq et al. recommended using peer reviews of API code in addition to API usability studies to uncover API related bugs and incorporate feedback~\cite{Farooq:2010:AUP:1753326.1753677}. In our work, instead of measuring overall API usability, we have focused specifically on the relationship of REST API usability with respect to usage examples.

\subsubsection{Controlled Studies}
Nasehi et al. performed a controlled study to understand if API unit tests can be used to provide as usage examples to facilitate API client developers \cite{5609553}. The researchers found that the examples from the unit tests helped understanding the API concepts better but it was challenging for the participants to locate relevant examples from the corpus of unit tests. Endrikat performed a controlled study with four groups of participants to understand the impact of API documentation on APIs that are implemented using programming languages with static and dynamic type systems \cite{Endrikat:2014:ADS:2568225.2568299}. Participants were given a set of failing unit tests and were asked to make them pass by writing code using the studied API. They found that the participant group using a statically typed API with explicit documentation could get the tests to pass in less time than the other groups. Ellis et al. performed a controlled study with two groups of developers to understand the impact of using a constructor vs. a factory method design pattern on API client development time \cite{Ellis:2007:FPA:1248820.1248863}. They found that API client developers needed more time to complete API tasks using factory method compared to using a constructor. Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned API related controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of a local API; and, the participants are professional software engineers.
